---
title: "R Notebook"
output: html_notebook
---

## 0. setup

In this analysis, tidyverse was widely used as a basic framework

```{r}
library(tidyverse)
```


### 0.1 read all required files

```{r}
## train
subject_train <- read_table("../input/UCI HAR Dataset/train/subject_train.txt", col_names = c("subject"))
X_train <- read_table("../input/UCI HAR Dataset/train/X_train.txt", col_names = FALSE)
y_train <- read_table("../input/UCI HAR Dataset/train/y_train.txt", col_names = c("activityId"))

# test
subject_test <- read_table("../input/UCI HAR Dataset/test/subject_test.txt", col_names = c("subject"))
X_test <- read_table("../input/UCI HAR Dataset/test/X_test.txt", col_names = FALSE)
y_test <- read_table("../input/UCI HAR Dataset/test/y_test.txt", col_names = c("activityId"))


features <- read_table2("../input/UCI HAR Dataset/features.txt", col_names = c("no", "feature"))
```


```{r}
activity_labels <- read_table("../input/UCI HAR Dataset/activity_labels.txt", col_names = c("activityId", "activityName"))
activity_labels
```


## 1. Merges the training and the test sets to create one data set.

```{r}
train <- bind_cols(subject_train, y_train, X_train)
test <- bind_cols(subject_test, y_test, X_test)
merged <- bind_rows(train, test)
```

Check the merged dataframe

```{r}
head(merged)
```

### 1.1 rename headers for analysis convenience 

```{r}
## add newname for the headers
features <- features %>% 
  mutate(
    newname = str_replace_all(feature, "-|,", "_"), ## replace '-', ',' to '_'
    newname = str_replace_all(newname, "\\(|\\)", "")  ## remove '(', ')'
  )

head(features)
```

```{r}
## renames headers usning features dataframe
colnames(merged)[3:ncol(merged)] <- features$newname

head(merged)
```


## 2. Extracts only the measurements on the mean and standard deviation for each measurement.

The dataframe was extracted from merged dataframe

```{r}
extracted <- merged %>% 
  select(contains(c("subject", "activityId", "mean", "std")))

colnames(extracted)
```

```{r}
head(extracted)
```
## 3. Uses descriptive activity names to name the activities in the data set

```{r}
activity_labels
```

```{r}
extracted
```

```{r}
extracted <-
  extracted %>% 
  left_join(activity_labels) %>% 
  relocate(subject, activityId, activityName)

head(extracted)
```


## 4. Appropriately labels the data set with descriptive variable names.

All header variables are replaced with descriptive variable names in section 1.1


## 5.From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
extracted_longer <-
  extracted %>% 
  pivot_longer(-c(subject, activityId, activityName), 
               names_to = "feature",
               values_to = "average")
extracted_longer
```





