---
title: "Ranking of national weather events affecting population and economic damage"
author: "Gyeong-Geun LEE"
date: '2020-12-03'
output:
  pdf_document: default
  html_document: default
---

# Synopsis

This analysis summarizes the weather events that occurred in the U.S. from 1966 to 2011, resulting in population health and property damage.
10 Major weather events were listed and the percentage were summarized. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE)
options(max.print = 50)
```


# Data processing

Basic data processing framework is loaded.

```{r, message=FALSE}
library(tidyverse)
library(patchwork)
```


The DB was downloaded and converted into csv file.

```{r, cache=TRUE}
db_link <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
dbfile <- tempfile()
download.file(db_link, dbfile)
raw <- read.csv(dbfile)
unlink(dbfile)
```


All names in the column are converted to lowercase for convenience.

```{r}
df0 <- 
  raw %>% 
  rename_all(tolower)
head(df0)
```

The raw data was glimpsed in order to understand the rough structure of the data.

```{r}
glimpse(df0)
```


Important features were seleted to analyze.

* `state`: state
* `bgn_date`" begin date
* `evtype`: event type
* `fatalities`
* `injuries`
* `propdmg`: property damage
* `propdmgexp`: exponent of property damage
* `cropdmg`: crop damage
* `cropdmgexp`: exponent of crop damage

```{r}
# selected features
df1 <- 
  df0 %>% 
  select(
    state,
    bgn_date,
    evtype, 
    fatalities, injuries, 
    propdmg, propdmgexp, cropdmg, cropdmgexp
  ) %>% 
  mutate(event_date = lubridate::mdy_hms(bgn_date)) %>% 
  relocate(evtype, state, event_date) %>% 
  select(-bgn_date)
```

In order to cut the limit of the year, the event number was shown according to year.

```{r}
df1 %>% 
  mutate(year = lubridate::year(event_date)) %>% 
  count(year) %>% 
  ggplot(aes(year, n)) +
  geom_point() +
  labs(x = "# of events")
```


Since the previous data is less reliable, only data after 1990 were considered.

```{r}
df1 <- 
  df1 %>% 
  filter(lubridate::year(event_date) >= 1990)
```


The NA data were removed.

```{r}
sum(!complete.cases(df1))
```


The exponent of property damage and crop damage were anlayzed

```{r}
table(df1$propdmgexp)
```


```{r}
table(df1$cropdmgexp)
```


The exponent was converted into numbers, and incoporated to mantissa.

```{r}
convert_exp <- function(s) {
  #s <- as.character()
  r <-
    case_when(
      s %in% c("", "-", "?", "+", "0", "h", "H") ~ "0",
      s %in% c("k", "K") ~ "3",
      s %in% c("m", "M") ~ "6",
      s %in% c("b", "B") ~ "9",
      s %in% c("1", "2", "3", "4", "5", "6", "7", "8") ~ s,
      TRUE ~ "0"
    )
  10^as.numeric(r)
}

df1 <-
  df1 %>% 
  mutate(
    propdmgexp2 = convert_exp(propdmgexp),
    cropdmgexp2 = convert_exp(cropdmgexp),
    prop = propdmg * propdmgexp2,           # finally converted value
    crop = cropdmg * cropdmgexp2            # finally converted value
  ) %>% 
  relocate(evtype:injuries, prop, crop)
df1
```


# Results

## events vs population health

The total was summarized for each event type.

```{r}
df1_population <-
  df1 %>% 
  group_by(evtype) %>% 
  summarize(
    fatalities = sum(fatalities),
    injuries = sum(injuries),
    .groups = "drop_last"
  )
df1_population
```


The fatalities and injuries were expressed in a figure.

```{r, fig.width=8, fig.height=4}
g1 <-
  df1_population %>%
  mutate(prop = fatalities / sum(fatalities)) %>% 
  arrange(desc(fatalities)) %>% 
  slice(1:10) %>%
  ggplot(aes(reorder(evtype, fatalities), fatalities)) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = paste(format(round(prop * 100, 1), nsmall = 1), "%")),
    hjust = -0.1, 
    size = 3
  ) +
  scale_y_continuous(limits = c(0, max(df1_population$fatalities)*1.2)) +
  coord_flip() +
  labs(x = "", y = "Fatalities")

g2 <-
  df1_population %>%
  mutate(prop = injuries / sum(injuries)) %>% 
  arrange(desc(injuries)) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(evtype, injuries), injuries)) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = paste(format(round(prop * 100, 1), nsmall = 1), "%")),
    hjust = -0.1, 
    size = 3
  ) +
  scale_y_continuous(limits = c(0, max(df1_population$injuries)*1.2)) +
  coord_flip() +
  labs(x = "", y = "Injuries")

g1 + g2 +
  plot_annotation(title = "10 Major weather events affecting population health")
```


## events vs economic consequences

```{r}
df1_economics <-
  df1 %>% 
  group_by(evtype) %>% 
  summarize(
    prop_sum = sum(prop),
    crop_sum = sum(crop),
    .groups = "drop_last"
  )
df1_economics
```


The population damage and crop damage were expressed in a figure.

```{r, fig.width=8, fig.height=4}
temp <- 
  df1_economics %>%
  mutate(
    dmg = prop_sum / 1e6,
    prop = dmg / sum(dmg)
  )
  
g1 <-
  temp %>% 
  arrange(desc(prop_sum)) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(evtype, dmg), dmg)) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = paste(format(round(prop * 100, 1), nsmall = 1), "%")),
    hjust = -0.1, 
    size = 3
  ) +
  scale_y_continuous(
    limits = c(0, max(temp$dmg)*1.2),
  ) +
  coord_flip() +
  labs(x = "", y = "Property damage $ (×1000)")

temp <- 
  df1_economics %>%
  mutate(
    dmg = crop_sum / 1e6,
    prop = dmg / sum(dmg)
  )
  
g2 <-
  temp %>%
  arrange(desc(crop_sum)) %>% 
  slice(1:10) %>% 
  ggplot(aes(reorder(evtype, dmg), dmg)) +
  geom_bar(stat = "identity") +
  geom_text(
    aes(label = paste(format(round(prop * 100, 1), nsmall = 1), "%")),
    hjust = -0.1, 
    size = 3
  ) +
  scale_y_continuous(
    limits = c(0, max(temp$dmg)*1.2),
  ) +
  coord_flip() +
  labs(x = "", y = "Crop damage $ (×1000)")

g1 + g2 +
  plot_annotation(title = "10 Major weather events affecting economic consequences")
```



