---
title: 'Part 1: Simulation Exercise Instructions'
author: "Gyeong-Geun LEE"
output:
  html_notebook: default
  pdf_document: default
---


# Overview

This report shows a simple example of the central limit theorem using an exponential distribution.


# Simultation

This simultation repeats 1000 times to calculate sample distrubition

```{r}
library(ggplot2)
#set.seed(131)
n <- 40        # no of samples
lambda <- 0.2  # rate of exponential distribution
nsim <- 10000   # no of simulation

rexp_matrix <- matrix(rexp(n * nsim, rate = lambda), nrow = nsim, ncol = n)
dim(rexp_matrix)
```

```{r}
sim_means <- rowMeans(rexp_matrix)
head(sim_means)
```


# Mean comparison

This is the mean of the samples means of my simulation results.

```{r}
mean_mns <- mean(sim_means)
mean_mns
```

This is theoretical eman of exponention distribution.

```{r}
theoretical_mean <- 1 / lambda
theoretical_mean
```

Two values are very similar. 

```{r}
ggplot(data = data.frame(mean = sim_means),  aes(mean)) +
  geom_histogram(aes(mean), color = "white", bins = 30) +
  geom_vline(xintercept = mean_mns, color = "blue") +
  geom_vline(xintercept = theoretical_mean, color = "red") 
```



# Variance comparision

This is the variance of the samples means of my simulation results.

```{r}
var_mns <- var(sim_means)
var_mns
```

This is theoretical variance of exponention distribution with n samples.

```{r}
theoretical_var <- (1 / (lambda^2)) / n
theoretical_var
```

Two values are very similar. 

This graph shows limit of 95% quantile of the mean +- 2 * sd.

```{r}
ggplot(data = data.frame(mean = sim_means),  aes(mean)) +
  geom_histogram(aes(mean), color = "white", bins = 30) +
  geom_vline(xintercept = mean_mns + 2 * sqrt(var_mns), color = "blue") +
  geom_vline(xintercept = mean_mns - 2 * sqrt(var_mns), color = "blue") +
  geom_vline(xintercept = theoretical_mean + 2 * sqrt(theoretical_var), color = "red") +
  geom_vline(xintercept = theoretical_mean - 2 * sqrt(theoretical_var), color = "red") 
```

# Distribution

The figure represents the Q-Q plot of samples means. There is a close similarity between sim_means and normal distribution represend with red line.

```{r}
qqnorm(sim_means)
qqline(sim_means, col = "red")
```






