---
title: "Part 2: Basic Inferential Data Analysis Instructions"
author: "Gyeong-Geun LEE"
output: html_notebook
---


# setup 

```{r}
library(tidyverse)   # load basic tools

df <- ToothGrowth    # load data
df
```

# EDA

The interaction between supp and dose was expressed as boxplots. As figures shows, the dose affects the length of tooth. In case low dose, Vitamin C (VC) act more effectively compared to Orange juice (OJ).

```{r}
df %>% 
  ggplot(aes(dose, len)) +
  geom_boxplot(aes(color = interaction(supp, dose))) +
  labs(x = "Dose", y = "Length of tooth") +
  theme(legend.title = element_blank())
```



# Basic summary of the data

Five numers of data was obtained, and sd was also calculated.

```{r}
df %>% 
  group_by(supp, dose) %>% 
  summarize(
    mean = mean(len),
    sd = sd(len),
    min = min(len),
    Q1 = quantile(len, 0.25),
    median = median(len),
    Q3 = quantile(len, 0.75),
    max = max(len),
    .groups = "drop"
  )
```


# Hypothesis tests

all conditions are splited.

```{r}
vc0.5 <- df %>% filter(supp == "VC", dose == 0.5) %>% pull(len)
vc1.0 <- df %>% filter(supp == "VC", dose == 1.0) %>% pull(len)
vc2.0 <- df %>% filter(supp == "VC", dose == 2.0) %>% pull(len)
oj0.5 <- df %>% filter(supp == "OJ", dose == 0.5) %>% pull(len)
oj1.0 <- df %>% filter(supp == "OJ", dose == 1.0) %>% pull(len)
oj2.0 <- df %>% filter(supp == "OJ", dose == 2.0) %>% pull(len)
```


## Comparision between supplement at the same dose

### dose = 0.5

The p-value is 0.005304, and null hypothesis was rejected.

```{r}
t.test(vc0.5, oj0.5, alternative = "two.sided", var.equal = TRUE)
```

### dose = 1.0

The p-value is 0.0007807, and null hypothesis is rejected.

```{r}
t.test(vc1.0, oj1.0, alternative = "two.sided", var.equal = TRUE)
```


### dose = 2.0

The p-value is 0.9637, and null hypothesis can not be rejected.

```{r}
t.test(vc2.0, oj2.0, alternative = "two.sided", var.equal = TRUE)
```


## Comparision between dose at the same supplement

### Vitamin C

Comparing dose 0.5 to 1.0, the p-value is 6.492e-07, and null hypothesis is rejected.

```{r}
t.test(vc0.5, vc1.0, alternative = "two.sided", var.equal = TRUE)
```


Comparing dose 1.0 to 2.0, the p-value is 3.398e-05, and null hypothesis is rejected.

```{r}
t.test(vc1.0, vc2.0, alternative = "two.sided", var.equal = TRUE)
```

### Orange juice

Comparing dose 0.5 to 1.0, the p-value is 8.358e-05, and null hypothesis is rejected.

```{r}
t.test(oj0.5, oj1.0, alternative = "two.sided", var.equal = TRUE)
```

Comparing dose 1.0 to 2.0, the p-value is 0.03736, and null hypothesis is rejected.

```{r}
t.test(oj1.0, oj2.0, alternative = "two.sided", var.equal = TRUE)
```


# Conclusion

1. Vitamin C is more effective than Orange juice to increase tooth length.
2. As the dose increases, the effect of supplements increases.

All data was assumed to be follow t-distribution.


















